A Multi-Modal Model for Prediction and Classification of Object Deformation during Robotic Manipulation

### 2016-04-13 twostagedo.tex
Changed illustrations for simulation of sponge.

The training of the model is performed offline using ground truthed data. The overall algorithm is given in \alref{alg:overall}. For clarity the algorithm is given in three phases: training, simulation (i.e. the prediction phase), and classification (i.e. matching prediction to new data). In testing the model new data is used, where this is typically new deformation actions on an object of known material and approximate shape. Having given an overview of the model we now relate our approach to those in the literature, before proceeding to describe the details of the spring-mass model of deformation (Section~\ref{sec:model}), and then the regression model of contact forces (Section~\ref{sec:forces}).

@@

For any preservation term, the potential energy $E$ is proportionally dependent on a preservation constraint $C$, where $k$ is the proportionality constant \eref{eq:EConstraint}.  The force $\vec{F}^i$ acting on the $i^{th}$ mass particle is defined as the negative gradient of this potential energy with respect to the position of the particle $\vec{p}_i$ \eref{eq:EForce}.
\begin{align}
 E(\vec{p}_1,...,\vec{p}_{n}) & = \frac{1}{2}kC(\vec{p}_1,...,\vec{p}_{n})^2, \label{eq:EConstraint} \\
 \vec{F}^i(\vec{p}_1,...,\vec{p}_{n}) & = -\frac{\partial}{\partial \vec{p}_i}E = -kC\frac{\partial C}{\partial \vec{p}_i}. \label{eq:EForce}
\end{align}
Which with damping becomes:
\begin{align}
 \label{eq:Fdamping}
 \vec{F}^i&(\vec{p}_1,...,\vec{p}_{n},\vec{v}_1,...,\vec{v}_{n}) = \nonumber \\
    &\left( -kC-k_d\sum_{1 \le j \le n} \dfrac{\partial C}{\partial \vec{p}_j}\vec{v}_j \right)
 \dfrac{\partial C}{\partial \vec{p}_i},
\end{align}
where $\vec{v}_j$ is the velocity of the $j^{th}$ particle connected to $i$, in a system with $n$ particles.

\subsubsection{Preservation of Length}
The constraint for the potential energy is given by:
\begin{equation}
 C = \dfrac{|\vec{p}_j-\vec{p}_i|-D_0}{D_0}, \label{eq:clength}
\end{equation}
where $\vec{p}_i$ and $\vec{p}_j$ are connected by a spring and $D_0$ is the rest length of that spring.  The force derived from this constraint pulls the masses in the direction of the spring that joins them [\fref{fig:forces}(a)].  Teschner normalised the value dividing by $D_0$ to make the elasticity constants scale independent.  Morris \cite{Morris2008} omitted that in his final equations for the force.  Here we use the normalised force $\vec{F}^i_D$ \eref{eq:FDlength}, which has the effect that, if we obtain a good set of parameters for a certain mesh of high resolution, the same set can be used for a mesh of lesser resolution that preserves the same type of symmetries.  Therefore, substituting \eref{eq:clength} in \eref{eq:Fdamping} we get:
\begin{align}
 \vec{F}^i_D(p_i,p_j,v_i,v_j) &= \dfrac{k_L}{D_0^2} \left( |\vec{p}_j-\vec{p}_i|-D_0 \right) \dfrac{\vec{p}_j-\vec{p}_i}{|\vec{p}_j-\vec{p}_i|}  + \nonumber \\
 & \dfrac{k_d}{D_0^2}\left( \dfrac{\vec{p}_j-\vec{p}_i}{|\vec{p}_j-\vec{p}_i|} \cdot (\vec{v}_j-\vec{v}_i) \right) \dfrac{\vec{p}_j-\vec{p}_i}{|\vec{p}_j-\vec{p}_i|} \label{eq:FDlength}
\end{align}
where $k_L$ is the linear spring constant and $k_d$ the damping constant.

@@

\begin{figure*}[!t]
\centering
\includegraphics[width=178mm]{arrio9}
\includegraphics[width=178mm]{arrio10}
\includegraphics[width=178mm]{arrio11}
\caption{Simulation of the sponge obtained with parameters $k_L=1898.22, k_d=1.24705, k_A=18629.2 k_{\phi}=764.649, max\_Force=183.71, yield=0.558793, creep=0.433517, max\_\alpha=0.73364$.  Using the customised integration scheme, the elastic-plastic model with all preservation terms, with the aid of geometric recovery routines and collision detection with the finger.  The block of material was represented by a mesh of $20 \times 10$ elements. Top: Training. Middle: Test 1. Bottom: Test 2}\label{fig:simulation_sponge}
\end{figure*}

\begin{figure*}[!t]
\centering
\includegraphics[width=178mm]{arrio12}
\includegraphics[width=178mm]{arrio13}
\includegraphics[width=178mm]{arrio14}
\caption{Simulation of the plasticine obtained with parameters $k_L=642.328, k_d=0.562201, k_A=29.2776 k_{\phi}=72.6202, max\_Force=765.912, yield=0.0790105, creep=0.664469, max\_\alpha=0.48842$.  Using the customised integration scheme, the elastic-plastic model with all preservation terms, with the aid of geometric recovery routines and collision detection with the finger.  The block of material was represented by a mesh of $20 \times 10$ elements. Top: Training. Middle: Test 1. Bottom: Test 2}\label{fig:simulation_plasticine}
\end{figure*}

>>>>

The training of the model is performed offline using ground truthed data. %The overall algorithm is given in \alref{alg:overall}.
For clarity the algorithm is given in three phases: training, simulation (i.e. the prediction phase), and classification (i.e. matching prediction to new data). In testing the model new data is used, where this is typically new deformation actions on an object of known material and approximate shape. Having given an overview of the model we now relate our approach to those in the literature, before proceeding to describe the details of the spring-mass model of deformation (Section~\ref{sec:model}), and then the regression model of contact forces (Section~\ref{sec:forces}).

@@

For every desired constraint $C(\vec{p}_1,...,\vec{p}_{n})$ acting over $n$ vertices, a potential energy $E$ is defined.  When the value of the constraint is zero, the potential energy is zero as well.  When the constraint is not satisfied, the potential energy increases.  This behaviour is obtained through the following definition for the energy:
\begin{align}
 E(\vec{p}_1,...,\vec{p}_{n}) & = \frac{1}{2}kC(\vec{p}_1,...,\vec{p}_{n})^2, \label{eq:EConstraint}
\end{align}
where $k$ is a proportionality constant \eref{eq:EConstraint} and $p_i$ is the position of particle $i$.

In order to reduce the energy of the system when a constraint is not satisfied, the force $\vec{F}^i$ acting on the $i^{th}$ mass particle is defined as the negative gradient of this potential energy with respect to the position of the particle $\vec{p}_i$ \eref{eq:EForce}.
\begin{align}
 \vec{F}^i(\vec{p}_1,...,\vec{p}_{n}) & = -\frac{\partial}{\partial \vec{p}_i}E = -kC\frac{\partial C}{\partial \vec{p}_i}. \label{eq:EForce}
\end{align}
Given the oscilatory nature of these equations, damping can be introduced to stop the oscillations.  This is modelled with a force acting in the opposite direction to the velocity of the affected particles:
\begin{align}
 \label{eq:Fdamping}
 \vec{F}^i&(\vec{p}_1,...,\vec{p}_{n},\vec{v}_1,...,\vec{v}_{n}) = \nonumber \\
    &\left( -kC-k_d\sum_{1 \le j \le n} \dfrac{\partial C}{\partial \vec{p}_j}\vec{v}_j \right)
 \dfrac{\partial C}{\partial \vec{p}_i},
\end{align}
where $\vec{v}_j = \frac{\partial \vec{p}_j}{\partial t}$ is the velocity of the $j^{th}$ particle connected to $i$, when the constraint acts on $n$ particles.

\subsubsection{Preservation of Length}
The constraint for the potential energy is given by:
\begin{equation}
 C = \dfrac{|\vec{p}_j-\vec{p}_i|-D_0}{D_0}, \label{eq:clength}
\end{equation}
where $\vec{p}_i$ and $\vec{p}_j$ are connected by a spring and $D_0$ is the rest length of that spring.  The force derived from this constraint pulls the masses in the direction of the spring that joins them [\fref{fig:forces}(a)].  Teschner normalised the value dividing by $D_0$ to make the elasticity constants scale independent.  Morris \cite{Morris2008} omitted that in his final equations for the force.  Here we use the normalised force $\vec{F}^i_D$ \eref{eq:FDlength}, which has the effect that, if we obtain a good set of parameters for a certain mesh of high resolution, the same set can be used for a mesh of lesser resolution that preserves the same type of symmetries.  Therefore, substituting \eref{eq:clength} in \eref{eq:Fdamping}, after some manipulation we get:
\begin{align}
 \vec{F}^i_D(p_i,p_j,v_i,v_j) &= \dfrac{k_L}{D_0^2} \left( |\vec{p}_j-\vec{p}_i|-D_0 \right) \dfrac{\vec{p}_j-\vec{p}_i}{|\vec{p}_j-\vec{p}_i|}  + \nonumber \\
 & \dfrac{k_d}{D_0^2}\left( \dfrac{\vec{p}_j-\vec{p}_i}{|\vec{p}_j-\vec{p}_i|} \cdot (\vec{v}_j-\vec{v}_i) \right) \dfrac{\vec{p}_j-\vec{p}_i}{|\vec{p}_j-\vec{p}_i|} \label{eq:FDlength}
\end{align}
where $k_L$ is the linear spring constant and $k_d$ the damping constant.

@@

\begin{figure*}[!t]
\centering
\includegraphics[width=178mm]{arrio9}
\includegraphics[width=178mm]{arrio10}
\includegraphics[width=178mm]{arrio11}
\caption{Simulation of the sponge obtained with parameters $k_L=3034.48, k_d=0.000237, k_A=103.568 k_{\phi}=66.774, max\_Force=1218.81, yield=0.000434973, creep=0.00890343, max\_\alpha=0.45204$.  Using the customised integration scheme, the elastic-plastic model with all preservation terms, with the aid of geometric recovery routines and collision detection with the finger.  The block of material was represented by a mesh of $20 \times 10$ elements. Top: Training. Middle: Test 1. Bottom: Test 2}\label{fig:simulation_sponge}
\end{figure*}

\begin{figure*}[!t]
\centering
\includegraphics[width=178mm]{arrio12}
\includegraphics[width=178mm]{arrio13}
%\includegraphics[width=178mm]{arrio14}
\caption{Simulation of the plasticine obtained with parameters $k_L=642.328, k_d=0.562201, k_A=29.2776 k_{\phi}=72.6202, max\_Force=765.912, yield=0.0790105, creep=0.664469, max\_\alpha=0.48842$.  Using the customised integration scheme, the elastic-plastic model with all preservation terms, with the aid of geometric recovery routines and collision detection with the finger.  The block of material was represented by a mesh of $20 \times 10$ elements. Top: Training. Middle: Test 1. Bottom: Test 2}\label{fig:simulation_plasticine}
\end{figure*}
